---
title: "Calculating BA"
output: html_notebook
---
Previously we have inspected the data in Inspecting_data.rmd...

We start with preparing the data, so that we don't have to run data inspection file.

```{r}
options(repos = "http://nexus.ki.se/repository/cran.r-project.org")
.libPaths("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/shun_trial/lib")
source("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/shun_trial/function.R")

library(tidyverse)
library(flexsurv)
library(haven)
library(survival)
library(broom)
library(ggplot2)
library(patchwork)
library(grid)
library(gridExtra)
library(checkmate) 
library(htmlTable)
library(survey)
library(tidyr)
library(dplyr)
library(psych)
library(splines)
library(ggeffects)
library(patchwork)
library(tableone) 
library(gridExtra)
library(ggtext)
library(gt)
library(stringr)
library(corrplot)
library(lubridate)
```

```{r}
amoris <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data/StudyPop.sas7bdat")
```

```{r}
#select our 17 biomarkers
biomarkers <- c("S_Alb", "S_Krea", "TC", "TG", "S_K", "S_P", "Fe_maet", "S_Urea", "S_LD", "S_Ca", "fS_TIBC", "fS_Jaern", "S_Alp", "S_Urat", "S_FAMN", "S_Hapt", "fS_Gluk")

AMORIS <- amoris %>%
  group_by(Kon) %>%
  filter(
    if_else(
      Kon == 1, 
      across(all_of(biomarkers), ~ .x > mean(.x, na.rm = TRUE) - 5 * sd(.x, na.rm = TRUE) & .x < mean(.x, na.rm = TRUE) + 5 * sd(.x, na.rm = TRUE)),
      across(all_of(biomarkers), ~ .x > mean(.x, na.rm = TRUE) - 5 * sd(.x, na.rm = TRUE) & .x < mean(.x, na.rm = TRUE) + 5 * sd(.x, na.rm = TRUE))
    )
  ) %>%
  ungroup()

AMORIS <- AMORIS[complete.cases(AMORIS[, biomarkers]), ]

rm(biomarkers)
```


Changing colneames and adding status and time
```{r}
colnames(AMORIS)[colnames(AMORIS) == "Kon"] <- "gender"
colnames(AMORIS)[colnames(AMORIS) == "Alder"] <- "age"
colnames(AMORIS)[colnames(AMORIS) == "Id"] <- "sampleID"

AMORIS <- AMORIS %>%
  mutate(status = ifelse(is.na(DODSDATn), 0, 1)) %>%
  mutate(time = as.numeric(difftime(lastDate, firstDate, units = "weeks")) / 4.33)

negative_counts <- sum(AMORIS$time < 0)
print(paste("Number of negative survival times:", negative_counts))

# Filtering out negative times
AMORIS <- AMORIS %>% filter(time >= 0)
```
```{r}
rm(negative_counts)
```


=========================================================================================
HERE WE START WITH BIOAGE
=========================================================================================


Now we know which biomarkers to use in the BA analysis
```{r}
#over 0.1 separatley for men and women, S_Ca is rounded up
biomarkers = c("S_Alb", "S_Krea", "TC", "TG", "S_K", "S_Urea", "S_LD", "S_Ca", "S_Urat", "S_FAMN", "S_Hapt", "fS_Gluk") 
```

Making our test set 
```{r}
## Making our training set and testing set
# 18 825 by checking obs when reading in the data frame, but since they trained on 30-75 the actual number is 12535
set.seed(123) # Setting a seed for reproducibility
small_amoris_cohort <- AMORIS[sample(nrow(AMORIS), 12535), ]
dim(small_amoris_cohort) # Check the dimensions
# Get SampleID of the selected cohort
selected_ids <- small_amoris_cohort$sampleID
# Get the SampleID of the rest of the cohort
rest_ids <- setdiff(AMORIS$sampleID, selected_ids)
# Filter the original data frame using these IDs to get the_rest_of_amoris
the_rest_of_amoris <- AMORIS[AMORIS$sampleID %in% rest_ids, ]
# Check the dimensions of the new data frames
dim(small_amoris_cohort)
dim(the_rest_of_amoris)

rm(rest_ids)
```



KDM
```{r}
## KDM
train_f = kdm_calc(small_amoris_cohort %>%
                     filter(age >=  25 & age <= 75, gender ==2),
                   biomarkers, fit = NULL, s_ba2 = NULL)
train_m = kdm_calc(small_amoris_cohort %>%
                     filter(age >= 25 & age <= 75, gender ==1),
                   biomarkers, fit = NULL, s_ba2 = NULL)
kdm_f = kdm_calc(the_rest_of_amoris %>%
                   filter (age >= 25 & age <= 75, gender ==2), #age >= 25 & age <= 75
                 biomarkers,
                 fit = train_f$fit,
                 s_ba2 = train_f$fit$s_ba2)
kdm_m = kdm_calc(the_rest_of_amoris %>%
                   filter (age >= 25 & age <= 75, gender ==1), #age >= 25 & age <= 75
                 biomarkers,
                 fit = train_m$fit,
                 s_ba2 = train_m$fit$s_ba2)
to1po2_kdm = rbind(kdm_f$data,kdm_m$data)

rm(train_f)
rm(train_m)
rm(kdm_f)
rm(kdm_m)
```

Plotting kdm sep. just to check
```{r}
agevar = c("kdm")
label = c("Modified-KDM\nBiological Age")
plot_ba(to1po2_kdm, agevar, label)
```



Phenoage
```{r}
train = phenoage_calc(small_amoris_cohort, biomarkers)
#Project into the AMORIS_2
phenoage = phenoage_calc(the_rest_of_amoris, biomarkers, fit = train$fit)
#Extract phenoage dataset, trained on 1 projected on 2 = to1po2, previously data
to1po2_phe = (phenoage$data)

rm(phenoage)
rm(train)
```

Plotting phenoage separatley 
```{r}
agevar = c("phenoage")
label = c("Modified-Phe\nBiological Age")
plot_ba(to1po2_phe, agevar, label)
```
 
 HD 
```{r}
#using 20-30 year olds as a reference range

hd_f = hd_calc (the_rest_of_amoris %>%
                  filter (gender ==2),
                small_amoris_cohort %>%
                  filter (gender ==2 & age >= 20 & age <= 30),
                biomarkers)
hd_m = hd_calc (the_rest_of_amoris %>%
                  filter (gender ==1),
                small_amoris_cohort %>%
                  filter (gender ==1 & age >= 20 & age <= 30),
                biomarkers)

#Extract HD dataset
to1po2_hd = rbind(hd_f$data, hd_m$data)

rm(hd_f)
rm(hd_m)
```
 Merging
```{r}
to1po2 = left_join(the_rest_of_amoris, to1po2_hd[, c("sampleID", "hd", "hd_log")], by = "sampleID") %>%
  left_join(., to1po2_kdm[, c("sampleID", "kdm", "kdm_advance")], by = "sampleID") %>%
  left_join(., to1po2_phe[, c("sampleID","phenoage","phenoage_advance")], by = "sampleID") 
```
 
 
  Plotting CA vs BA
```{r}
agevar_ap = c("kdm","phenoage","hd","hd_log")
#prepare labels
label_ap = c("KDM amoris\nBiological Age",
             "phenoage amoris\nPhenotypic Age",
             "Homeostatic\nDysregulation",
             "Log\nHomeostatic\nDysregulation")
#plot age vs bioage
plot_ba(to1po2, agevar_ap, label_ap)
```

Residuals
```{r}
BA_residuals <- function(BA){
  data = to1po2 %>% drop_na(BA)
  # Basic model = regress on age alone
  model <- parse(text = sprintf("lm(%s ~ ns(age, df = 3), data = data)", BA)) %>% eval()
  model_predict <- ggpredict(model, terms = c("age"))
  data[,"BA_res"] <- NA
  data[!is.na(data[BA]),"BA_res"] <- resid(model)
  return(residuals(model))
}
for(BA in c("kdm", "phenoage", "hd", "hd_log")){
  BA_res <- paste0(BA, "_res")
  to1po2[,BA_res] = NA
  to1po2[!is.na(to1po2[BA]),BA_res] <- BA_residuals(BA)
}
rm(list=c("BA","BA_res"))
```

Correlation with BA, HAVE TO REDOOOOO THISSS!!!
```{r}
get_axis_type <- function(labels){
  return(rep("float", length(labels)) %>% setNames(names(labels))) # Create function to generate axis_type variables for BAA plots
}
plot_baa(data=to1po2, 
         agevar=names(c("kdm"="KDM",
                        "kdm_res"="KDM\nresidual",
                        "phenoage"="PhenoAge",
                        "phenoage_res"="PhenoAge\nresidual",
                        "hd_log"="HD (log)",
                        "age"="Chronological\nage")), 
         label=c("kdm"="KDM",
                 "kdm_res"="KDM\nresidual",
                 "phenoage"="PhenoAge",
                 "phenoage_res"="PhenoAge\nresidual",
                 "hd_log"="HD (log)",
                 "age"="Chronological\nage"), 
         axis_type=get_axis_type(c("kdm"="KDM",
                                   "kdm_res"="KDM\nresidual",
                                   "phenoage"="PhenoAge",
                                   "phenoage_res"="PhenoAge\nresidual",
                                   "hd_log"="HD (log)",
                                   "age"="Chronological\nage")))

```

##Table of summary statistics! 
Preparing varables
```{r}
# Calculate age at FirstDate in years
to1po2$AgeAtFirstDate <- as.numeric(difftime(to1po2$firstDate, to1po2$FODDATn, units = "days") / 365.25)
# Calculate age at LastDate in years
to1po2$AgeAtLastDate <- as.numeric(difftime(to1po2$lastDate, to1po2$FODDATn, units = "days") / 365.25)
```

```{r}
SveDem <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data/SveDem.sas7bdat")
SveDem_format <- read_sas("W:/C6_AmorisGuest/Sara Hagg/Karolina Gustavsson/Data/SveDem_fomats.sas7bdat")
```

```{r}
all_codes <- c("304", "290", "290A", "F00", "305", "290B", "G30", "311A", "G308", "G309", "G301", "F009", "F002", "G300", "F001", "F000", "293", "290E", "F01", "293,1", "F019", "F012", "F018", "F013", "F011", "F010", "F0183", "306", "290X", "F02", "290W", "F03", "294B", "G311", "311B", "G318A", "311C", "F051", "311X", "F039", "F020", "F024", "F028", "290", "290,1", "F022", "F021", "F023", "F03-P")
alzheimers_codes <- c("304", "290", "290A", "F00", "305", "290B", "G30", "311A", "G308", "G309", "G301", "F009", "F002", "G300", "F001", "F000")
vascular_codes <- c("293", "290E", "F01", "293,1", "F019", "F012", "F018", "F013", "F011", "F010", "F0183")
other_codes <- c("306", "290X", "F02", "290W", "F03", "294B", "G311", "311B", "G318A", "311C", "F051", "311X", "F039", "F020", "F024", "F028", "290", "290,1", "F022", "F021", "F023", "F03-P")
unclear_type <- c("293,9","293,4", "293,3", "293,5", "F0010", "F0013", "F0180") #these are not in SveDem

#All cases
to1po2$ADRD <- ifelse(to1po2$Dementia == "", 0, 1)
to1po2$Alzheimers <- ifelse(to1po2$Dementia %in% alzheimers_codes, 1, 0)
to1po2$Vascular <- ifelse(to1po2$Dementia %in% vascular_codes, 1, 0)
to1po2$Other <- ifelse(to1po2$Dementia %in% other_codes, 1, 0)
to1po2$Unclear <- ifelse(to1po2$Dementia %in% unclear_type, 1, 0)

#Svedem unique
to1po2$S_ADRD <- ifelse(to1po2$Dementia %in% all_codes & to1po2$sampleID %in% SveDem$Id, 1, 0)
to1po2$S_Alzheimers <- ifelse(to1po2$Dementia %in% alzheimers_codes & to1po2$sampleID %in% SveDem$Id, 1, 0)
to1po2$S_Vascular <- ifelse(to1po2$Dementia %in% vascular_codes & to1po2$sampleID %in% SveDem$Id, 1, 0)
to1po2$S_other <- ifelse(to1po2$Dementia %in% other_codes & to1po2$sampleID %in% SveDem$Id, 1, 0)

#Not linked to SveDem, from other registers
to1po2$NS_ADRD <- ifelse(to1po2$Dementia %in% all_codes & !to1po2$sampleID %in% SveDem$Id, 1, 0)
to1po2$NS_Alzheimers <- ifelse(to1po2$Dementia %in% alzheimers_codes & !to1po2$sampleID %in% SveDem$Id, 1, 0)
to1po2$NS_Vascular <- ifelse(to1po2$Dementia %in% vascular_codes & !to1po2$sampleID %in% SveDem$Id, 1, 0)
to1po2$NS_other <- ifelse(to1po2$Dementia %in% other_codes & !to1po2$sampleID %in% SveDem$Id, 1, 0)
```


Table with overview of cohort
```{r}
# Load necessary libraries
library(dplyr)
library(broom)

# Function to calculate descriptive statistics with mean ± SD for selected columns 
# and count for specific columns
calculate_descriptive_stats <- function(df, column, include_sd = FALSE, use_log_p_value = FALSE, count_only = FALSE) {
  overall <- df
  males <- df %>% filter(gender == '1')
  females <- df %>% filter(gender == '2')

  if (count_only || column == 'status') {
    # Count the number of cases and calculate the percentage
    count_overall <- sum(overall[[column]], na.rm = TRUE)
    count_males <- sum(males[[column]], na.rm = TRUE)
    count_females <- sum(females[[column]], na.rm = TRUE)
    
    percentage_overall <- (count_overall / nrow(overall)) * 100
    percentage_males <- (count_males / nrow(males)) * 100
    percentage_females <- (count_females / nrow(females)) * 100
    
    return(data.frame(Mean_Overall = paste(count_overall, "(", round(percentage_overall, 2), "%)", sep = ""),
                      Mean_Males = paste(count_males, "(", round(percentage_males, 2), "%)", sep = ""),
                      Mean_Females = paste(count_females, "(", round(percentage_females, 2), "%)", sep = ""),
                      P_Value = NA))
  }

  # Calculation for other columns
  mean_overall <- mean(overall[[column]], na.rm = TRUE)
  mean_male <- mean(males[[column]], na.rm = TRUE)
  mean_female <- mean(females[[column]], na.rm = TRUE)

  if (include_sd) {
    sd_overall <- sd(overall[[column]], na.rm = TRUE)
    sd_male <- sd(males[[column]], na.rm = TRUE)
    sd_female <- sd(females[[column]], na.rm = TRUE)
    mean_sd_overall <- paste(round(mean_overall, 2), "±", round(sd_overall, 2))
    mean_sd_male <- paste(round(mean_male, 2), "±", round(sd_male, 2))
    mean_sd_female <- paste(round(mean_female, 2), "±", round(sd_female, 2))
  } else {
    mean_sd_overall <- as.character(round(mean_overall, 2))
    mean_sd_male <- as.character(round(mean_male, 2))
    mean_sd_female <- as.character(round(mean_female, 2))
  }

  # Perform t-test
  p_column <- ifelse(use_log_p_value, "TG_log", column)
  if (column %in% names(df)) {
    t_test <- tidy(t.test(males[[p_column]], females[[p_column]], na.rm = TRUE))
    p_value <- t_test$p.value
  } else {
    p_value <- NA  # Assign NA if the column doesn't exist
  }

  return(data.frame(Mean_Overall = mean_sd_overall, Mean_Males = mean_sd_male, 
                    Mean_Females = mean_sd_female, P_Value = p_value))
}

# Add log-transformed TG to the dataset
to1po2$TG_log <- log(to1po2$TG)

# Define columns for analysis, including new columns
columns_to_analyze <- c('age', 'AgeAtLastDate', 'status', 'kdm', 'kdm_res', 'phenoage', 
                        'phenoage_res', 'hd', 'hd_log', 'hd_log_res', 'S_Krea', 'TC', 'TG', 'fS_Gluk', 'fS_Jaern', 
                        'S_Ca', 'S_Urat', 'fS_TIBC', 'Fe_maet', 'S_Alb', 'S_Alp', 'S_K', 'S_FAMN', 
                        'S_LD', 'S_Hapt', 'S_Urea', 'S_P', 'TimeUntilFollowUp', 
                        'ADRD', 'Alzheimers', 'Vascular', 'Other', 'Unclear', 
                        'S_ADRD', 'S_Alzheimers', 'S_Vascular', 'S_other',
                        'NS_ADRD', 'NS_Alzheimers', 'NS_Vascular', 'NS_other') # Added new columns

# Columns for which mean ± SD is required
columns_sd_required <- c('age', 'AgeAtLastDate', 'kdm', 'kdm_res', 'phenoage', 'phenoage_res', 
                         'hd', 'hd_log', 'hd_log_res', 'S_Krea', 'TC', 'TG', 'fS_Gluk', 'fS_Jaern', 'S_Ca', 
                         'S_Urat', 'fS_TIBC', 'Fe_maet', 'S_Alb', 'S_Alp', 'S_K', 'S_FAMN', 'S_LD', 'S_Hapt', 
                         'S_Urea', 'S_P', 'TimeUntilFollowUp')

# Columns for which count and percentage are required
columns_count_required <- c('ADRD', 'Alzheimers', 'Vascular', 'Other', 'Unclear', 
                            'S_ADRD', 'S_Alzheimers', 'S_Vascular', 'S_other', 'status',
                            'NS_ADRD', 'NS_Alzheimers', 'NS_Vascular', 'NS_other') # Include 'status' and new columns

# Create an empty DataFrame for results
results_df <- data.frame(Measure = character(), Mean_Overall = character(), 
                         Mean_Males = character(), Mean_Females = character(), 
                         P_Value = numeric(), stringsAsFactors = FALSE)

# Calculate stats for each column
for (column in columns_to_analyze) {
  include_sd <- column %in% columns_sd_required
  use_log_p_value <- column == 'TG'
  count_only <- column %in% columns_count_required
  stats <- calculate_descriptive_stats(to1po2, column, include_sd, use_log_p_value, count_only)
  stats$Measure <- column
  results_df <- rbind(results_df, stats)
}

# Renaming measures in the results DataFrame
measure_names <- c('Chronological age', 'Age at follow up', 'Mortality at censorship', 'kdm', 'kdm_res', 
                   'phenoage', 'phenoage_res', 'hd', 'hd_log', 'hd_log_res', 'S_Krea', 'TC', 'TG', 'fS_Gluk', 
                   'fS_Jaern', 'S_Ca', 'S_Urat', 'fS_TIBC', 'Fe_maet', 'S_Alb', 'S_Alp', 'S_K', 'S_FAMN', 'S_LD', 
                   'S_Hapt', 'S_Urea', 'S_P', 'Time until follow up', 'ADRD', 'Alzheimers', 'Vascular', 'Other', 
                   'Unclear', 'S_ADRD', 'S_Alzheimers', 'S_Vascular', 'S_other', 'NS_ADRD', 'NS_Alzheimers', 'NS_Vascular', 'NS_other')
results_df$Measure <- factor(results_df$Measure, levels = columns_to_analyze, labels = measure_names)

# Add N values as a separate row
n_values <- data.frame(Measure = "N", 
                       Mean_Overall = as.character(nrow(to1po2)),
                       Mean_Males = as.character(sum(to1po2$gender == '1')),
                       Mean_Females = as.character(sum(to1po2$gender == '2')),
                       P_Value = NA)

results_df <- rbind(n_values, results_df)

# Print the results DataFrame
print(results_df)


```

